---
title: "whichmuesum"
output: html_document
---
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(xml2)
```
# Source Overview
Base Page 
```{r} 
base_url <- "https://whichmuseum.com/place/united-states-2682"
base_url
``` 
According to the WhichMuesum United States page, there are 12046 museum in United States. 
# Visual Categories
From the website sidebar, the main museum catergories are
```{r}
whichmuseum_categories <- tibble(
  category = c(
    "History & Anthropology",
    "Science & Technology",
    "Nature & Natural History",
    "Art & Design",
    "Specialized & Alternative"
  ),
  count = c(7517,1954,1821,1517,812)
)
whichmuseum_categories
``` 
```{r}
whichmuseum_categories %>%
  ggplot(aes(x = reorder(category,count), y = count)) +
  geom_col()+
  coord_flip()+
  labs(
    title = "WhichMuseum Categories",
    x = "Categories",
    y = "Number of Museum"
  )
``` 
#Robot.txt check 
Robot.txt page
```{r}
robot_url <- "https://whichmuseum.com/robots.txt"
robot_url
``` 
Summary: 
This robots.txt file does not disallow general crawling for `User-agent: *`. It blocks pages related to users, logins, comments, image licensing, and contact information, but allows the crawling of standard pages.
Important disallowed path include
```{r}
disallowed_path<-tibble(
  path=c(
     "/claim-request",
    "/user.json",
    "/login?",
    "/signup?",
    "/write-review",
    "/img-license/",
    "/user/",
    "/review/translate/",
    "/museum-text-license/",
    "/logout",
    "/museum-widget/",
    "/contact?"
  )
)
disallowed_path
```
# Example museum page 
Example museum page:
```{r}
example_url <- "https://whichmuseum.com/museum/fort-stevens-state-park-warrenton-13187"
example_url
```
```{r}
example_page <- read_html(example_url)
example_text <- example_page %>%
  html_text2()
cat(example_text)
```
#Links on Example page
```{r}
example_links <- example_page %>%
  html_elements("a")

example_links <- tibble(
  text = html_text2(example_links),
  href = html_attr(example_links, "href")
)

example_links %>%
  filter(!is.na(href)) %>%
  print(n = 50)
```
#Field visible in WhichMuseum page
museum pages appear to contain
```{r}
visible_fields <- tibble(
  field = c(
    "Museum name",
    "Rating",
    "Number of ratings",
    "National ranking",
    "State ranking",
    "Short description",
    "Category",
    "Address",
    "City",
    "State",
    "Country",
    "Phone number",
    "Website link",
    "Map link"
  ),
  useful_for_database = c(
    TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE,
    TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE
  )
)
visible_fields
```
