---
title: "dataset_exploration"
format: html
editor: visual
---

```{r}
# Load packages
library(tidyverse)
library(tidygeocoder)
library(janitor)
library(tidytext)
library(leaflet)
library(lubridate)
```

## Street art !

```{r}
# Reading the now-comma-free headers + the rest and cleaning
murals <- read_csv("../data/Murals/streetartcities_US_2026-06.csv",
                   na = c("", "NA", "Unknown", "Untitled")) %>% 
  mutate(across(32:ncol(murals), as.character)) %>% #otherwise it tries to parse them as lgl
  clean_names() #%>% # From janior package :)
```

```{r}
# Cleaning/wrangling further
murals_clean <- murals %>% 
  #select(where(function(x) !all(is.na(x)))) # could also do the ~ shorthand
  remove_empty("cols") %>% # From janitor package
  select(-country) %>% # It's US for all here
  select(1:attribute_last_seen) %>% # removing the rest of the cols since most are almost 100% NA and not really helpful for me
  mutate(description_lower = tolower(description)) # for text analysis? or do later
```

```{r}
# Using stringr+regex to fix the weird formatting
for(i in 1:ncol(murals_clean)){
  if(is.character(murals_clean[[i]])){
    murals_clean[[i]] <- str_remove_all(murals_clean[[i]], "<.*?>|\\[|\\]|\"")
  }
}
```

remove the photo created + date situation? unsure... -'photo taken' or 'curated by' -  (more html tags too?) -project -mural -NA -art - the place... - numbers

```{r}
# Text analysis of description
remove_words <- tibble(
  word = c(NA, "nbsp", "project", "mural", "murals", "painted", "paint", "painting", "street", "NA", "art", "https", "artist", "artists", "san", "www.instagram.com", "instagram", "created", "located", "curated")
)

murals_description_words <- murals_clean %>% 
  mutate(description_lower = str_remove_all(description_lower, "\\d+")) %>% # removing nums (like years, etc)
  unnest_tokens(output = word, input = description_lower) %>% 
  anti_join(get_stopwords(source = "stopwords-iso"), by="word") %>% #could change the source...
  anti_join(remove_words, by = "word") %>% 
  group_by(city_id) %>% 
  count(word) %>% 
  slice_max(n, n = 3)
```

## Historic Districts

#\| eval: false

```{r}

# Loading historic landmarks data for historic districts!
historic_landmarks <- read_csv("../data/Historic Districts/historic_landmarks.csv") # .. means one folder up
```

```{r}

# Filtering the historic landmarks dataset to make sure that only historic districts are present
historic_districts <- historic_landmarks %>% 
  clean_names() %>% 
  mutate(listed_date = mdy(listed_date),
         nhl_designated_date = mdy(nhl_designated_date),
         across(where(is.character) & -c(prefix, street_number, external_link, other_names), str_to_title),
         district_name = property_name) %>%
  filter(str_detect(district_name, "Historic District")) %>% 
  mutate(district_name_lower = str_remove(district_name, " Historic District")) %>% 
  mutate(address = paste(property_name, county, state, sep=" "))
```

Note: `category of property` variable -- district for most, but site for some... but it is all caps sometimes and not some other times ex: `Fort Union Trading Post National Historic Site Historic District (Boundary Increase)` -- marked as `site` and it's not really a historic district in the way I was thinking...

```{r}

# CLeaning the historic_districts dataset so only the columns we want are present


# and adding location
historic_districts <- historic_districts %>%
  geocode(address, method = 'arcgis', lat = latitude , long = longitude)

#method = census resulted in NAs
```

then i did `write.csv(historic_districts, "historic_districts_clean.csv")` so we wouldn't have to re-run the above over and over again (since the geocoding with arcgis is kinda slow). so you can just run the below

## Historic districts w/o the above!

```{r}
historic_districts <- read_csv("../data/Historic Districts/historic_districts_clean.csv")

# Mapping w/ leaflet!
m = leaflet(data = historic_districts) %>%
   addTiles() %>%
   addCircleMarkers(~longitude, ~latitude, popup = ~property_name)
m
```

## News outlets

```{r}
# Loading news outlet data
outlets <- read_csv("../data/News Outlets/us-news-deserts-news-source-export-06-12-12-140252_SOURCE_UNC.csv")
```
