---
title: "historic_districts_exploration"
format: html
editor: visual
---

```{r}
# Load packages
library(tidyverse)
library(tidygeocoder)
library(janitor)
library(tidytext)
library(leaflet)
library(tigris)
library(readr)
library(sf)
```

```{r}
# Cleaning the historic_districts dataset so only the columns we want are present (add later)
```

--

```{r}
unemployment_data_usda <- read_csv("../data/EK_general/Unemployment2023.csv")
```
```{r}
unemployment_wider <- unemployment_data_usda %>% 
  clean_names() %>% 
  mutate(
    #separate("\d{4}")
    year = str_sub(attribute, -4, str_length(attribute)),
    attribute = str_sub(attribute, 1, -6) %>% 
      str_to_lower() %>% 
      str_replace_all("_", " ")
  )
```
```{r}
# now actually pivoting wider
unemployment_wider <- unemployment_wider %>% 
  pivot_wider(
    names_from = attribute,
    values_from = value
  )
```





```{r}
library(devtools)
install_github("mikeasilva/blsAPI")
```
```{r}
library(blsAPI)

```
```{r}
library(BLSloadR)
bls_data <- get_laus(
  geography = "county",
  monthly_only = FALSE,
  transform = TRUE,
  suppress_warnings = TRUE,
  return_diagnostics = FALSE,
  #cache = check_bls_cache_env()
)
```
```{r}
write_csv(bls_data, file = "raw_bls_data_1")
```
```{r}
bls_data %>% 
  summarize(max_val = max(year))
```

```{r}
bls_data_2 <- bls_data %>% 
  filter(year == 2025,
         measure_text == "unemployment rate",
         period == "M13") # just annual
```



```{r}
write_csv(bls_data_2, file = "bls_data.csv")
```
```{r}
bls_data_3  <- read_csv("../data/EK_general/bls_data.csv") %>% 
  mutate(NAMELSAD = str_remove(area_text, ",.*"), #before comma
         STUSPS = str_trim(str_replace(area_text, "^.*,",""))) #after comma
# using stringr to mutate to match shapefile for easier joining
```

```{r}
# Get counties shape files
counties_sf <- tigris::counties(cb = TRUE) %>% 
  st_transform(crs = 4326)
```
```{r}
# join data to shapefile
mapping_data <- counties_sf %>% 
  left_join(bls_data_3, by = c("NAMELSAD", "STUSPS"))

```
^ value is the unemployment

```{r}
pal <- colorNumeric(
  palette = "Spectral",           #also tried piyg and ylor and didn't love either
  domain = mapping_data$value
)



leaflet(mapping_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(
    fillColor = ~pal(value),
    fillOpacity = .75,
    color = "white", # border color
    weight = 1,
    smoothFactor = 0.5, # slightly crisper borders -- default is 1 (higher values > more simplification > jaggier borders but shorter rendering)
    # add the highlight/hover and tooltip things
  ) %>% 
  addLegend(
    pal = pal,
    value = mapping_data$value, # same as values   = ~total_num_districts
    position = "bottomright",
    title = "Legend"
  )
```


```{r}

states_sf <- tigris::states(cb = TRUE, resolution = "20m") %>% 
  st_transform(crs = 4326)

# Join data to shapefile
choropleth_area_data <- states_sf %>% 
  left_join(by_state, by = c("NAME" = "state")) %>% 
  left_join(areas, by = c("NAME" = "state_or_territory")) %>% 
  mutate(total_acreage_hd = total_acreage) %>% 
  select(-total_acreage)
```


```{r}
# Import cleaned (sort of) dataset
historic_districts <- read_csv("../data/Historic Districts/historic_districts_clean3_specific_addresses.csv")

# Mapping w/ leaflet!
prelim_map <- leaflet(data = historic_districts) %>%
   addTiles() %>%
   #addProviderTiles("Stadia.StamenWatercolor") %>% 
   addCircleMarkers(~longitude, ~latitude, popup = ~property_name, radius = 3)
prelim_map
```

```{r}

by_state <- historic_districts %>% 
  group_by(state) %>% 
  summarise(total_acreage = sum(acreage_of_property, na.rm=TRUE),
            total_num_districts = n(),
            across(25:last_col(), ~ sum(.x, na.rm = TRUE))) # NEED TO ADD THE COUNTS FOR EACH CATEGORY

categories_counts <- by_state %>% 
  select(state, 4:ncol(by_state)) %>% 
  column_to_rownames(var = "state") %>%   # Promotes state from a regular column to R row name
  t() %>% # Transposes (rows become columns, columns become rows) (returns a matrix)
  as.data.frame() %>% 
  rownames_to_column(var = "category") %>% 
  mutate(usa_total = rowSums(across(where(is.numeric)), na.rm = TRUE),
         category_og = category, 
         category = gsub("_", " ", str_remove(category, "aos_")),
         category_nice = str_to_title(category)) %>% 
  bind_rows( # so that we can calculate percentages in the future!!
    summarise(., 
              across(where(is.numeric), base::sum), 
              across(where(is.character), ~"Total"))
  )
```

```{r}
#print(df_total)
```


```{r}
# Import (mostly) cleaned dataset
historic_districts <- read_csv("../data/Historic Districts/historic_districts_clean4.csv")

# Mapping w/ leaflet!
prelim_map <- leaflet(data = historic_districts) %>%
   addTiles() %>%
   #addProviderTiles("Stadia.StamenWatercolor") %>% 
   addCircleMarkers(~longitude, ~latitude, popup = ~property_name, radius = 3)
prelim_map
```

```{r}

# Mapping w/ leaflet!
prelim_map2 <- leaflet(data = historic_districts) %>%
   addTiles() %>%
   addMarkers(clusterOptions = markerClusterOptions(), popup = ~property_name)
  #addCircleMarkers(~longitude, ~latitude, popup = ~property_name, radius = 3)
prelim_map2
```

## Choropleths!!!

```{r}
# Group by state for data from historic_districts

by_state <- historic_districts %>% 
  group_by(state) %>% 
  summarise(total_acreage = sum(acreage_of_property, na.rm=TRUE),
            total_num_districts = n())

```

```{r}
by_state <- historic_districts %>% 
  group_by(state) %>% 
  summarise(total_acreage = sum(acreage_of_property, na.rm=TRUE),
            total_num_districts = n(),
            across(25:last_col(), \(x) sum(x, na.rm = TRUE)))
```

```         
```

```{r}
temp_df <- as.data.frame(t(by_state[, 25:ncol(by_state)])) %>% 
  rename(counts = V1) %>% 
  tibble::rownames_to_column(var = "category")
```

```{r}
temp_df <- by_state %>% 
  select(state, 4:ncol(by_state)) %>% 
  column_to_rownames(var = "state") %>%   
  # ^ Promotes state from a regular column to R row name
  t() %>% # Transposes (rows become columns, columns become rows) (returns a matrix)
  as.data.frame() %>% 
  rownames_to_column(var = "category")
```

```{r}
areas <- read_csv("../data/Historic Districts/us_areas_cleaned.csv")
```

cb = TRUE simplfies the file, resolution = 20m is the lowest, this gives 2024 data

```{r}
# Get state geometries
states_sf <- tigris::states(cb = TRUE, resolution = "20m")
```

```{r}
# Join data to shapefile
map_data <- states_sf %>% 
  left_join(by_state, by = c("NAME" = "state")) %>% 
  left_join(areas, by = c("NAME" = "state_or_territory")) %>% 
  mutate(total_acreage_hd = total_acreage) %>% 
  select(-total_acreage)
```

```{r}
# Standardizing
# historic district acreage by total state land acreage
map_data <- map_data %>% 
  mutate(standardized_hd_acreage = total_acreage_hd/land_area_acres) 
```

## Standardized historic district area by state

```{r}
# Create a color palette for the map
# might want to change this to binning or quantile or something

my_palette <- colorNumeric(
  palette = "viridis", 
  domain = map_data$standardized_hd_acreage,
  na.color = "transparent"
)
```

```{r}
leaflet(map_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(
    fillColor = ~my_palette(standardized_hd_acreage),
    fillOpacity = .75,
    color = "white", # border color
    weight = 1,
    smoothFactor = 0.5, # slightly crisper borders -- default is 1 (higher values > more simplification > jaggier borders but shorter rendering)
    # add the highlight/hover and tooltip things
  ) %>% 
  addLegend(
    pal = my_palette,
    value = map_data$standardized_hd_acreage, # same as values   = ~total_num_districts
    position = "bottomright",
    title = "Hist. Dist. area/state land area"
  )
```

## Land area by state

Chloropleth map for \# of acres of districts \# MAYBE SHOULD JOIN DATA FOR THE TOTAL ACREAGE OF THE WHOLE STATE TO STANDARDIZE \# \^ for this should find the area of each state and then divide the acreage by that...

```{r}
# Create a color palette for the map
# might want to change this to binning or quantile or something

my_palette <- colorNumeric(
  palette = "viridis", 
  domain = map_data$land_area_acres,
  na.color = "transparent"
)
```

```{r}
leaflet(map_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(
    fillColor = ~my_palette(land_area_acres),
    fillOpacity = .75,
    color = "white", # border color
    weight = 1,
    smoothFactor = 0.5, # slightly crisper borders -- default is 1 (higher values > more simplification > jaggier borders but shorter rendering)
    # add the highlight/hover and tooltip things
  ) %>% 
  addLegend(
    pal = my_palette,
    value = map_data$land_area_acres, # same as values   = ~total_num_districts
    position = "bottomright",
    title = "land area by state"
  )
```

```{r}
# 
# acreage_of_property na.rm = TRUE --> ignore missing values

```

```{r}
# choosing the var

#explore_var <- map_data$total_num_districts
```

```{r}
# Create a color palette for the map
# might want to change this to binning or quantile or something

my_palette <- colorNumeric(
  palette = "viridis", 
  domain = map_data$total_num_districts,
  na.color = "transparent"
)
```

```{r}
leaflet(map_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(
    fillColor = ~my_palette(total_num_districts),
    fillOpacity = .75,
    color = "white", # border color
    weight = 1,
    smoothFactor = 0.5, # slightly crisper borders -- default is 1 (higher values > more simplification > jaggier borders but shorter rendering)
    # add the highlight/hover and tooltip things
  ) %>% 
  addLegend(
    pal = my_palette,
    value = map_data$total_num_districts, # same as values   = ~total_num_districts
    position = "bottomright",
    title = "# of historic districts"
  )
```

```{r}
# Create a color palette for the map
# might want to change this to binning or quantile or something

my_palette <- colorNumeric(
  palette = "viridis", 
  domain = map_data$total_acreage_hd,
  na.color = "transparent"
)
```

```{r}
leaflet(map_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(
    fillColor = ~my_palette(total_acreage_hd),
    fillOpacity = .75,
    color = "white", # border color
    weight = 1,
    smoothFactor = 0.5, # slightly crisper borders -- default is 1 (higher values > more simplification > jaggier borders but shorter rendering)
    # add the highlight/hover and tooltip things
  ) %>% 
  addLegend(
    pal = my_palette,
    value = map_data$total_acreage_hd, # same as values   = ~total_num_districts
    position = "bottomright",
    title = "acreage of historic districts"
  )
```
