---
title: "murals_exploration"
format: html
editor: visual
---

```{r}
# Load packages
library(tidyverse)
library(tidygeocoder)
library(janitor)
library(tidytext)
library(leaflet)
library(lubridate)
library(arcgisgeocode)

library(tigris)
library(sf)
library(ggplot2)
library(dplyr)
```

# Street art !

## Importing data / cleaning / wrangling

```{r}
# Reading the now-comma-free headers + the rest and cleaning
murals <- read_csv("../data/Murals/streetartcities_US_2026-06.csv",
                   na = c("", "NA", "Unknown", "Untitled"))

murals <- murals %>% 
  mutate(across(32:ncol(murals), as.character)) %>% #otherwise it tries to parse them as lgl
  clean_names()
```

```{r}
# Cleaning/wrangling further
murals_clean <- murals %>% 
  #select(where(function(x) !all(is.na(x)))) # could also do the ~ shorthand
  remove_empty("cols") %>% # From janitor package
  select(-country) %>% # It's US for all here
  select(1:attribute_last_seen) %>% 
  # ^ removing the rest of the cols since most are almost 100% NA and not really helpful for me
  mutate(description_lower = tolower(description)) # for text analysis? or do later
```

```{r}
# Using stringr+regex to fix the weird formatting
for(i in 1:ncol(murals_clean)){
  if(is.character(murals_clean[[i]])){
    murals_clean[[i]] <- str_remove_all(murals_clean[[i]], "<.*?>|\\[|\\]|\"")
  }
}
```

### Getting regions

```{r}
# Cache shapefiles so you don't have to re-download them
options(tigris_use_cache = TRUE)

# Download cartographic boundary polygons for US Census Regions
region_polygons <- regions() %>% 
  st_transform(crs = 4326) # ensuring the same coordinate reference systems
```

```{r}
# Lil chart of regions
ggplot() +
  geom_sf(data = region_polygons, aes(fill = NAME), color = "white")
```

```{r}
# https://sunsp.net/rspatial_book/spatialops.html 2.5
# Automatically filters murals inside polygons AND pulls over the "NAME" column
murals_clean <- st_as_sf(murals_clean, coords = c("longitude", "latitude"), crs = 4326)
murals_clean <- st_join(murals_clean, region_polygons, join = st_within) %>%
  #filter(!is.na(NAME)) %>%       # Optional: removes murals that didn't fall into any region
  rename(region = NAME) %>% 
  select(-c(31, 32, 33, 35, 36, 37, 38))

# st_intersects could theoretically also work but it would include on the borders
# st_contains would work if the two dfs were switched (bc a point can't contain a polygon)
```

## Text analysis

### Tokenizing

```{r}
# Text analysis of description
# Just part 1 -- tokenizing

# Used claude for some help with the custom stopwords
custom_stop_words <- c(
  # Standard English
  "the", "of", "and", "a", "in", "to", "by", "is", "for", "on",
  "this", "s", "with", "that", "it", "was", "from", "as", "at",
  "i", "an", "are", "one", "has", "you", "all", "he", "my", "be",
  "their", "we", "who", "can", "her", "but", "out", "our", "they",
  "which", "so", "into", "have", "had", "an", "its", "been", "not",
  "also", "more", "up", "or", "about", "than", "when", "will",
  
  # Domain-specific (high-frequency, low signal)
  "mural", "murals", "art", "artist", "artists", "painted", "street",
  "wall", "walls", "photo", "work", "project", "city", 
  #"community",
  "building", "piece", "created", "located", "curated", "paint",
  "arts", "side", "part", "new", 
  #"local", 
  "life", 
  #"people",
  "through", "impermanent", # "impermanent" is included as domain-specific — it appeared 469 times, almost certainly from a boilerplate template phrase rather than meaningful content.
  
  # URL / web noise
  "https", "http", "com", "www", "instagram", "www.instagram.com",
  
  # Emily's words
  NA, "nbsp", "project", "painting", "street", "NA", "san", "amp", "photos"
  
  # The places
  # Add/figure out later
  
)

# To use with tidytext:
 custom_df <- tibble(word = custom_stop_words, lexicon = "custom")
 stop_words_combined <- bind_rows(stop_words, custom_df)

murals_descrip_words <- murals_clean %>% 
  #mutate(description_lower = str_remove_all(description_lower, "\\d+")) %>% # removing nums (like years, etc)
  unnest_tokens(output = word, input = description_lower) %>% 
  #anti_join(get_stopwords(source = "stopwords-iso"), by="word") %>% #could change the source...
  anti_join(stop_words_combined, by = "word")
```

### Prev

```{r}
# IDEAS:
# region, state
# conservative vs liberal wrods/polarizing words
# west coast
# size of city

# Prev:
#murals_descrip_words %>% 
#  group_by(city_id) %>% 
#  count(word) %>% 
#  slice_max(n, n = 3)
```

### By region

```{r}
words_by_region <- murals_descrip_words %>% 
  group_by(region) %>% 
  count(word)
```

```{r}
words_by_region %>% 
  slice_max(n, n = 50)
```

### Further TA

- https://cran.r-project.org/web/packages/conversim/vignettes/Analyzing_Similarities_between_Two_Long_Speeches.html
- https://www.r-bloggers.com/2015/10/text-mining-analysis-some-theory-and-practice-in-r/ \^ https://www.rdatamining.com/home
- https://medium.com/@celine.vdr/natural-language-processing-word-embeddings-e0b2edc773d2 \^ https://github.com/Celinevdr/Analysing-Similarities-and-Differences-Between-Republicans-and-Democrats-Through-Word-Embeddings/blob/master/GitHub_Finalproject.Rmd

my - https://medium.com/data-science/r-packages-for-text-analysis-ad8d86684adb - https://stackoverflow.com/questions/8898521/finding-2-3-word-phrases-using-r-tm-package - https://www.rdatamining.com/examples/text-mining - https://www.slideshare.net/slideshow/r-by-example-mining-twitter-for/8504941

vids - https://www.youtube.com/watch?v=7cwBhWYHgsA&t=44s - https://campus.datacamp.com/courses/introduction-to-natural-language-processing-in-r/representations-of-text?ex=12

specifically cosine similarity

#### Based on medium article

https://medium.com/@celine.vdr/natural-language-processing-word-embeddings-e0b2edc773d2
https://github.com/Celinevdr/Analysing-Similarities-and-Differences-Between-Republicans-and-Democrats-Through-Word-Embeddings/blob/master/GitHub_Finalproject.Rmd

### Topic modelling
I don't think this really shows anything...
```{r}
# Preprocessing
library(SnowballC) # for wordstem function 
# this chunk took forever
descrip_words_df <- murals_descrip_words %>% 
  mutate(word = wordStem(word)) %>% 
  count(id, word, sort = TRUE) %>% 
  ungroup()
```
```{r}
library(topicmodels) # for lda
library(reshape2) # dependency of topicmodels
# Building DTM
descrip_words_dtm <- descrip_words_df %>% 
  cast_dtm(id, word, n) # cols = words, and cell vals = counts

# ^ tidytext equivalent of the text2vec approach from earlier.

# Fitting LDA
lda <- LDA(descrip_words_dtm, k=5, control = list(seed = 1234)) 
# ^ seed thingy makes it reproducible. 1234 is arbitrary

# Extracting top terms
topics <- tidy(lda, matrix = "beta") 
# ^ Extracts the beta matrix (per-topic word probabilities.) Each row is a word-topic pair with a beta value representing how probable that word is for that topic. Higher beta = more characteristic of that topic.

# Get just top 10
top_terms <- topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
# Plot
top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```

-----

# Most frequent usage of words by region?
```{r}
descrip_words_df_2 <- murals_descrip_words %>% 
  mutate(word = wordStem(word)) %>% 
  count(region, word, sort = TRUE)
```

```{r}
# a graph for that

descrip_words_df_2 %>% 
  arrange(desc(n)) %>% 
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  # ^ convert word col to a factor with levels set in reverse order of their current appearance
  group_by(region) %>% 
  slice_max(order_by = n, n = 5) %>% #top_n(5) %>% 
  ungroup() %>% 
  ggplot(aes(word, n, fill = region)) +
  geom_col() +
  labs(x = NULL, y = "n") +
  #facet_wrap(~region, ncol = 5, scales = "free") +
  coord_flip()
```


### Brief cosine similarity
```{r}
library(quanteda)

# Creating a corpus
murals_corpus <- corpus(murals_clean, text_field = "description")
#docnames(murals_corpus) <- docid

# Make document feature matrix
dfmat <- dfm(tokens(murals_corpus)) #murals_descrip_words
```
```{r}

```



Before I dive into the analysis, the following code is useful to identify already the similarities and differences between party platforms, using a simple ```quanteda```function and testing similarity using the cosine measure of similarity. The closer the cosine score is to 1, the more similar are the documents. 

The table below shows the documents with the lowest cosine score and as such, the most different.

```{r echo=FALSE}
cosine_sim <- textstat_simil(dfmat, method = "cosine", margin = "documents")
results <- (as.matrix(cosine_sim))
pairs <- subset(melt(results), value!=1)
pairs <- data.frame(t(apply(pairs, 1, sort)))
pairs <- unique(pairs)
pairs <- pairs [ ,c(3,2,1)]
names(pairs)<- c("Platform1", "Platform2", "Cosine Similarity")
pairs <- pairs[order(pairs$`Cosine Similarity`),] 
head(pairs, n=10)
```

The table below shows the most similar party platforms, with similarity scores closest to 1. 

```{r echo=FALSE}
pairs <- pairs[order(pairs$`Cosine Similarity`, decreasing = TRUE),] 
head(pairs, n=10)
```

### actual analysis

```{r}
library(text2vec)

# Setting up

# Fitting the model: GloVe
## Constructing a document-term matrix

### For the whole dataset
complete_tokens <- itoken(murals_clean$description_lower,
                          preprocessor = tolower, 
                          # ^ lowercases all inputs... already done (CBL!)
                          tokenizer = word_tokenizer, 
                          # ^ splits text into individual word tokens
                          ids = murals_clean$id, # maybe change this since
                          # she does row number for id
                          progressbar = FALSE)
                          # ^ supress consoloe output
```
```{r}
# Table of each unique word
vocab_all <- create_vocabulary(complete_tokens)
```
```{r}
# Turning the vocabulary into a vectorizer function (basically a lookup that maps each word to a column index). This is what lets create_dtm() know which column each word belongs to.
vectorizer_all <- vocab_vectorizer(vocab_all)
```
```{r}
# Building the document-term matrix — a sparse matrix where rows are documents (indexed by id) and columns are vocabulary words, with cell values representing word counts. Sparse format means zeroes aren't stored explicitly, keeping memory usage low.
dtm = create_dtm(complete_tokens, vectorizer_all)
```
```{r}
# Should i do one for each region??? not sure...
```
```{r}
# Setting up GloVe algorithm



vocab_all = prune_vocabulary(vocab_all, term_count_min = 5L)
vectorizer_all = vocab_vectorizer(vocab_all)
tcm = create_tcm(platforms_tokens, vectorizer_all, skip_grams_window = 5L)
glove = GlobalVectors$new(word_vectors_size = 50, vocabulary = vocab_all, x_max = 10)
wv_main = glove$fit_transform(tcm, n_iter = 50, convergence_tol = 0.00001)
wv_context = glove$components
word_vectors =  wv_main + t(wv_context)
```



